Towards atomistic resolution structure of phosphatidylcholine glycerol backbone and choline headgroup at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}

### 2015-03-19 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Contributions & Acknowledgements of Matti Javanainen and Joona Tynkkynen

\author{Alexandru Botan}
%\affiliation{The authors are listed in alphaphetical order.}
%\affiliation{The author list is not completed.}
\affiliation[Lyon CNRS]{Institut Lumi\`ere Mati\`ere, UMR5306 Universit\'e Lyon 1-CNRS, Universit\'e de Lyon 69622 Villeurbanne, France}
%
%\author{Andrea Catte}
%\affiliation{University of East Anglia, Norwich, United Kingdom}
\author{Fernando Favela-Rosales}
\affiliation[Mexico]{Departamento de F{\'i}sica - CINVESTAV}
\author{Patrick Fuchs}
\affiliation[CNRS paris]{Institut Jacques Monod, CNRS, Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, Paris, France}
\alsoaffiliation[Diderot University]
{Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, Paris, France}
\author{Matti Javanainen}
\affiliation[Tampere University of Technology]
{Tampere University of Technology}
\author{Waldemar Kulig}
\affiliation[Tampere University of Technology]
{Department of Physics, Tampere University of Technology, Tampere, Finland}
\author{Antti Lamberg}
\affiliation[Kyoto University]
{Department of Chemical Engineering, Kyoto University, Kyoto, Japan}
\author{Markus S. Miettinen}
\affiliation[Freie Universitat Berlin] 
{Fachbereich Physik, Freie Universitat Berlin, Berlin, Germany}
\author{Luca Monticelli}
\affiliation[UMR] 
{5IBCP, CNRS UMR 5086, Lyon, France}
\author{Jukka Maatta}
\affiliation[Aalto University]
{Aalto University, Espoo, Finland}
\author{O. H. Samuli Ollila} 
\email{samuli.ollila@aalto.fi.}
\affiliation[Aalto University]
{Aalto University, Espoo, Finland}
\author{Marius Retegan}
\affiliation[Max Planck]
{Max Planck Institute for Chemical Energy Conversion, Mulheim an der Ruhr, Germany}
\author{Tomasz Rog}
\affiliation[Tampere University of Technology]
{Tampere University of Technology, Tampere, Finland}
\author{Hubert Santuz}
\affiliation[INSERM]
{INSERM, U1134, DSIMB; Institut National de la Transfusion Sanguine (INTS); Laboratoire d'Excellence GR-Ex, Paris, France}
\alsoaffiliation[Diderot]
{Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, Paris, France}
\author{Joona Tynkkynen}
\affiliation[Tampere University of Technology]
{Tampere University of Technology, Tampere, Finland}

@@

Slipid \cite{jambeck12}          & DPPC & 128 & 3840 & 323 & 150 & 100 & [\citenum{slipidsFILES}]$^*$ & SI \\
Slipid \cite{jambeck12b}          & POPC & 128 & 5120 & 303 & 200 & 150 & [\citenum{slipidsFILESpopc}]$^*$ & SI \\

@@

None of the studied force fields fulfils these criteria completely, however CHARMM36 is pretty close. 
This is not surprising since the dihedral potentials in this model are tuned to reproduce these parameters better against experiments~\cite{klauda10}.
The next models in the list are CHARMM36-UA~\cite{henin08,lee14} and H\"ogberg08~\cite{hogberg08} which is also not surprising since
these models are using CHARMM bonded potentials for glycerol backbone and choline. The fourth and the fifth models in the list, MacRog~\cite{maciejewski14} and
GAFFlipid~\cite{dickson12}, have independently determined dihedral potentials. All the models based on Gromos potentials and Slipid perform less well.
In the present work we subject the CHARMM36, MacRog, GAFFlipid and Berger-POPC-07 to a more careful comparison including the sterospecific labeling  
(Fig.~\ref{HGorderparameters2}), and atomistic level structure and responses to the dehydration and cholesterol content in the following sections.
These models are selected for more detailed studies since they are the best representatives of different dihedral potential parametrization techniques 
(CHARMM36, MacRog, GAFFlipid), and the Berger based models are the most used lipid model in the literature.

@@

@@

\subsection{Author Contributions}
\todo{Please write a text summarizing your contribution} \\
{\it Alexandru Botan} \\
{\it Fernando Favela}  \\
{\it Patrick Fuchs} \\
{\it Matti Javanainen} \\
{\it Waldemar Kulig} prepared the MD simulations with cholesterol for the MacRog FF \\
{\it Antti Lamberg}  \\
{\it Markus S. Miettinen}  \\
{\it Luca Monticelli}  \\
{\it Jukka Maatta}  \\
{\it O. H. Samuli Ollila} Designed and managed the work. Ran and analyzed several simulations. Wrote the manuscript.  \\
{\it Marius Retegan}  \\
{\it Tomasz Rog}  \\
{\it Hubert Santuz}  \\
{\it Joona Tynkkynen}  \\

>>>>

\author{Alexandru Botan}
%\affiliation{The authors are listed in alphaphetical order.}
%\affiliation{The author list is not completed.}
\affiliation[Lyon CNRS]{Institut Lumi\`ere Mati\`ere, UMR5306 Universit\'e Lyon 1-CNRS, Universit\'e de Lyon 69622 Villeurbanne, France}
%
%\author{Andrea Catte}
%\affiliation{University of East Anglia, Norwich, United Kingdom}
\author{Fernando Favela-Rosales}
\affiliation[Mexico]{Departamento de F{\'i}sica - CINVESTAV}
\author{Patrick Fuchs}
\affiliation[CNRS paris]{Institut Jacques Monod, CNRS, Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, Paris, France}
\alsoaffiliation[Diderot University]
{Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, Paris, France}
\author{Matti Javanainen}
\affiliation[Tampere University of Technology]
{Department of Physics, Tampere University of Technology, Tampere, Finland}
\author{Waldemar Kulig}
\affiliation[Tampere University of Technology]
{Department of Physics, Tampere University of Technology, Tampere, Finland}
\author{Antti Lamberg}
\affiliation[Kyoto University]
{Department of Chemical Engineering, Kyoto University, Kyoto, Japan}
\author{Markus S. Miettinen}
\affiliation[Freie Universitat Berlin] 
{Fachbereich Physik, Freie Universitat Berlin, Berlin, Germany}
\author{Luca Monticelli}
\affiliation[UMR] 
{5IBCP, CNRS UMR 5086, Lyon, France}
\author{Jukka Maatta}
\affiliation[Aalto University]
{Aalto University, Espoo, Finland}
\author{O. H. Samuli Ollila} 
\email{samuli.ollila@aalto.fi.}
\affiliation[Aalto University]
{Aalto University, Espoo, Finland}
\author{Marius Retegan}
\affiliation[Max Planck]
{Max Planck Institute for Chemical Energy Conversion, Mulheim an der Ruhr, Germany}
\author{Tomasz Rog}
\affiliation[Tampere University of Technology]
{Department of Physics, Tampere University of Technology, Tampere, Finland}
\author{Hubert Santuz}
\affiliation[INSERM]
{INSERM, U1134, DSIMB; Institut National de la Transfusion Sanguine (INTS); Laboratoire d'Excellence GR-Ex, Paris, France}
\alsoaffiliation[Diderot]
{Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, Paris, France}
\author{Joona Tynkkynen}
\affiliation[Tampere University of Technology]
{Department of Physics, Tampere University of Technology, Tampere, Finland}

@@

Slipids \cite{jambeck12}          & DPPC & 128 & 3840 & 323 & 150 & 100 & [\citenum{slipidsFILES}]$^*$ & SI \\
Slipids \cite{jambeck12b}          & POPC & 128 & 5120 & 303 & 200 & 150 & [\citenum{slipidsFILESpopc}]$^*$ & SI \\

@@

None of the studied force fields fulfils these criteria completely, however CHARMM36 is pretty close. 
This is not surprising since the dihedral potentials in this model are tuned to reproduce these parameters better against experiments~\cite{klauda10}.
The next models in the list are CHARMM36-UA~\cite{henin08,lee14} and H\"ogberg08~\cite{hogberg08} which is also not surprising since
these models are using CHARMM bonded potentials for glycerol backbone and choline. The fourth and the fifth models in the list, MacRog~\cite{maciejewski14} and
GAFFlipid~\cite{dickson12}, have independently determined dihedral potentials. All the models based on Gromos potentials and Slipids perform less well.
In the present work we subject the CHARMM36, MacRog, GAFFlipid and Berger-POPC-07 to a more careful comparison including the sterospecific labeling  
(Fig.~\ref{HGorderparameters2}), and atomistic level structure and responses to the dehydration and cholesterol content in the following sections.
These models are selected for more detailed studies since they are the best representatives of different dihedral potential parametrization techniques 
(CHARMM36, MacRog, GAFFlipid), and the Berger based models are the most used lipid model in the literature.

@@

MJ and JT acknowledge and CSC - IT Center for Science for computational resources (project number tty3979).
MJ also acknowledges the Finnish Doctoral Programme in Computational Sciences (FICS) for funding.

@@

\subsection{Author Contributions}
\todo{Please write a text summarizing your contribution} \\
{\it Alexandru Botan} \\
{\it Fernando Favela}  \\
{\it Patrick Fuchs} \\
{\it Matti Javanainen} prepared and performed simulations with multiple lipid models and analyzed the results. Supervised the work of JT.\\
{\it Waldemar Kulig} prepared the MD simulations with cholesterol for the MacRog FF \\
{\it Antti Lamberg}  \\
{\it Markus S. Miettinen}  \\
{\it Luca Monticelli}  \\
{\it Jukka Maatta}  \\
{\it O. H. Samuli Ollila} Designed and managed the work. Ran and analyzed several simulations. Wrote the manuscript.  \\
{\it Marius Retegan}  \\
{\it Tomasz Rog}  \\
{\it Hubert Santuz}  \\
{\it Joona Tynkkynen} prepared and performed the dehydration simulations with the MacRog FF.\\
